---
title: "Online Study Data Cleaning"
subtitle: "Further exclusions"
author: "Carina Fan"  
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = "../data/")
```

This file runs further exclusions (e.g., for health conditions) after an initial clean for data validity.

# Set up

```{r packages}
# load packages
library(psych)
library(magrittr)
library(tidyverse)
```

```{r today}
today = "2023-07-26"
```

```{r data}
df = paste0(today, "-demo-clean.csv") %>% 
  read.csv(stringsAsFactors = FALSE)
```


<!-- ======================================================================= -->


# Exclude based on health history and PHQ

```{r}
# names of health variables to exclude based on
healthNames = df %>% 
  select(starts_with("health_")) %>% 
  select(-c("health_other", "health_other_text", "health_alcdrug")) %>% 
  names()

# number of participants endorsing each condition
df.health = df %>% 
  select(all_of(healthNames)) %>% 
  mutate_all(as.character)
sapply(df.health, function(x) sum(str_count(x, "Yes"), na.rm = TRUE))

# column indices of variables to exclude 
healthCols = which(names(df) %in% healthNames)

# initialize vector of subjects with neuro conditions 
healthSubjects = vector(mode = "character")

## if subject responded Yes to any neuro condition, 
## add their row number to neuroSubject
for (row in 1:nrow(df)) {
  for (col in 1:length(healthCols)) {
    if (df[row, healthCols[col]] == "Yes") {
      healthSubjects = append(healthSubjects, row)
    }
  }
}

# number of subjects to exclude based on health history
length(unique(healthSubjects)) 

# number of subjects to exclude based on PHQ (scores >= 20)
# df$phq_total %<>% as.character() %>% as.numeric()
# df %>%
#   filter(phq_total >= 20) %>%
#   nrow()

# exclude subjects 
df.clean = df %>% 
  filter(!row_number() %in% healthSubjects)
         # phq_total < 20) 
df.excluded = df |> 
  filter(row_number() %in% healthSubjects)

# number of participants after all exclusions
nrow(df.clean)
```


<!-- ======================================================================= -->


# Demographics of participants in clean data

```{r}
# age & edu
df.clean %>% 
  select(age, edu) %>% 
  describe()

# gender
df.clean %>% 
  group_by(gender) %>% 
  summarise(n = n())
```


<!-- ======================================================================= -->


# Demographics of previously excluded participants

The following are numbers after excluding for health/neurological history, but retaining participants with high PHQ scores or substance abuse.

```{r}
df.clean |> 
  filter(phq_total >= 20) |>
  group_by(sdam) |> 
  summarise(n = n())
```

Of participants with PHQ scores above threshold, `r (580 / (580 + 139) * 100) |> round(1)`% had SDAM. Note that these participants are not excluded from the extraclean data.

```{r}
df.clean |> 
  filter(health_alcdrug == "Yes") |> 
  group_by(sdam) |> 
  summarise(n = n())
```

Of participants with alcohol/drug abuse, `r (609 / (609 + 306) * 100) |> round(1)`% had SDAM.

```{r}
df.clean |> 
  filter(phq_total >= 20 | health_alcdrug == "Yes") |> 
  group_by(sdam) |> 
  summarise(n = n())
```

Of those with PHQ or substance abuse, `r (1129 / (1129 + 420) * 100) |> round(1)`% had SDAM.

<!-- ======================================================================= -->


# Export data

```{r}
# df.clean %>% 
#   write.csv(paste0(today, "-demo-extraclean.csv"), row.names = FALSE)
```


<!-- ======================================================================= -->


# Session info

```{r session_info, include = TRUE}
sessionInfo()
```