---
title: "On the relationship between episodic autobiographical memory and spatial navigation"
subtitle: "Analysis script"
author: "Carina Fan, Hervé Abdi, & Brian Levine"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output: 
  github_document:
    toc: true
---

This file contains the code and output from all the analyses reported in the main text and supplemental materials of the manuscript.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Set up

```{r packages}
# load packages
library(corrplot)
library(ExPosition)
library(gplots)
library(gridExtra)
library(lmSupport)
library(psych)
# devtools::install_github("https://github.com/HerveAbdi/PTCA4CATA")
library(PTCA4CATA)
library(readxl)
library(reshape)
library(TInPosition)
library(wrapr)
library(magrittr)
library(tidyverse)
```

```{r colours}
# colour codes for plots 
colourCodeicor = colorRampPalette(c("blue", "yellow", "red"))(n = 299)
colourCode.sam = c("deepskyblue", "darkolivegreen", "firebrick3", "darkorchid3")
colourCode.osiq = c("darkgoldenrod", "slateblue")
colourCode.bfi = c("orange1", "lightsalmon1", "navajowhite3", "gold1", "peachpuff3")
```

```{r load_data}
# load data
df.raw = read.csv("../../data/old/2019-10-03-demo-raw.csv")
df.clean = read.csv("../../data/old/2019-10-03-demo-clean.csv", na.strings = "")

# number of responses recorded
nrow(df.raw)

# number of responses after preliminary cleaning
nrow(df.clean)
```

## Exclusions

```{r exclusions}
# names of health variables to exclude based on
neuroNames = df.clean %>% 
  select(starts_with("health_")) %>% 
  select(-c("health_other", "health_other_text")) %>% 
  names()

# number of participants endorsing each condition
df.neuro = df.clean %>% select(neuroNames)
df.neuro %<>% mutate_all(as.character)
sapply(df.neuro, function(x) sum(str_count(x, "Yes"), na.rm = TRUE))

# column indices of variables to exclude 
neuroIndices = which(names(df.clean) %in% neuroNames)

# initialize vector of subjects with neuro conditions 
neuroSubjects = vector(mode = "character")

## if subject responded Yes to any neuro condition, 
## add their row number to neuroSubject
for (row in 1:nrow(df.clean)) {
  for (col in 1:length(neuroIndices)) {
    if (is.na(df.clean[row, neuroIndices[col]])) {
      next
    }
   else if (df.clean[row, neuroIndices[col]] == "Yes") {
      neuroSubjects = append(neuroSubjects, row)
    }
  }
}

# number of subjects to exclude based on neurological history
length(unique(neuroSubjects)) 

# number of subjects to exclude based on PHQ (scores >= 20)
df.clean$phq_total %<>% as.character() %>% as.numeric()

df.clean %>%
  filter(phq_total >= 20) %>%
  nrow()

# exclude subjects based on health conditions
df = df.clean %>% filter(!row_number() %in% neuroSubjects) 

# exclude subjects based on PHQ
df %<>% filter(phq_total < 20)

# number of participants after all exclusions
nrow(df)
```

<!-- ======================================================================= -->

# Demographics

Age and education:

```{r age_edu}
df %>% 
  summarise(age_mean = mean(age),
            age_sd = sd(age),
            edu_mean = mean(edu),
            edu_sd = sd(edu))

# age range
range(df$age)
```

Gender:

```{r gender}
df %>% 
  group_by(gender) %>% 
  summarise(n = n())
```

<!-- ======================================================================= -->

# Factor structure of the Survey of Autobiographical Memory (SAM)

```{r setup_sam}
# select SAM items
df.sam = df %>%
  select(gender, starts_with("sam")) %>%
  select(-c("sam_epi", "sam_sem", "sam_spa", "sam_fut"))

# rename variables
names(df.sam) = c(
  "gender",
  paste0("E", seq(1, 8)),
  paste0("F", seq(1, 6)),
  paste0("M", seq(1, 6)),
  paste0("S", seq(1, 6))
)

# split dataframe by gender for supplemental analyses
df.sam.female = df.sam %>% filter(gender == "Female")
df.sam.male   = df.sam %>% filter(gender == "Male")

# remove gender column from dataframes
df.sam$gender        = NULL
df.sam.female$gender = NULL
df.sam.male$gender   = NULL

# assign colours to variables
colour.sam = as.vector(rep("", ncol(df.sam)))

for (i in 1:ncol(df.sam)) {
  switch(substr(names(df.sam)[i], 1, 1),
         E = {colour.sam[i] = colourCode.sam[1]},
         F = {colour.sam[i] = colourCode.sam[2]},
         M = {colour.sam[i] = colourCode.sam[3]},
         S = {colour.sam[i] = colourCode.sam[4]})
}

# convert to numeric
df.sam %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)
df.sam.female %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)
df.sam.male %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)
```

```{r pca_sam}
pca.sam = epPCA(
  df.sam,
  scale = FALSE,
  center = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
pca.sam$ExPosition.Data$t
```

<!-- ======================================================================= -->

# Relationship between SAM and Object-Spatial Imagery Questionnaire (OSIQ)

## Univariate analysis (linear regression)

```{r setup_sam_osiq}
# select SAM and OSIQ items
df.sam.osiq = df %>%
  select(starts_with("sam"),
         osiq_s_1, osiq_s_9, osiq_s_13, osiq_s_14, osiq_s_18, osiq_s_27r, osiq_s_29,
         osiq_o_4, osiq_o_12, osiq_o_17, osiq_o_22, osiq_o_25, osiq_o_28, osiq_o_30)

## convert to numeric
df.sam.osiq %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

# calculate OSIQ domain scores based on mini OSIQ items

## object
df.sam.osiq$osiq_mini_o =
  df.sam.osiq %>%
  select(starts_with("osiq_o")) %>%
  rowMeans()

## spatial
df.sam.osiq$osiq_mini_s =
  df.sam.osiq %>%
  select(starts_with("osiq_s")) %>%
  rowMeans()

# calculate SAM domain scores based on raw item scores

## episodic
df.sam.osiq$sam_epi_raw = df.sam.osiq %>%
  select(starts_with("sam_event")) %>%
  rowMeans()

## spatial
df.sam.osiq$sam_spa_raw = df.sam.osiq %>%
  select(starts_with("sam_spatial")) %>%
  rowMeans()

## semantic
df.sam.osiq$sam_sem_raw = df.sam.osiq %>%
  select(starts_with("sam_semantic")) %>%
  rowMeans()

## future
df.sam.osiq$sam_fut_raw = df.sam.osiq %>%
  select(starts_with("sam_future")) %>%
  rowMeans()
```

```{r regression_sam_osiq}
# SAM-episodic
model_samE_osiq = lm(sam_epi_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq)
model_samE_osiq %>% summary()
model_samE_osiq %>% modelEffectSizes()

# SAM-spatial
model_samS_osiq = lm(sam_spa_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq)
model_samS_osiq %>% summary()
model_samS_osiq %>% modelEffectSizes()

# SAM-future
model_samF_osiq = lm(sam_fut_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq)
model_samF_osiq %>% summary()
model_samF_osiq %>% modelEffectSizes()

# SAM-semantic
model_samM_osiq = lm(sam_sem_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq)
model_samM_osiq %>% summary()
model_samM_osiq %>% modelEffectSizes()
```

## Multivariate analysis (partial least squares correlation)

```{r setup_osiq}
# select mini OSIQ items
df.osiq = df %>%
  select(osiq_s_1, osiq_s_9, osiq_s_13, osiq_s_14, osiq_s_18, osiq_s_27r, osiq_s_29,
         osiq_o_4, osiq_o_12, osiq_o_17, osiq_o_22, osiq_o_25, osiq_o_28, osiq_o_30)

## convert to numeric
df.osiq %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

# rename variables
names(df.osiq) = c(
  "s1", "s9", "s13", "s14", "s18", "s27", "s29",
  "o4", "o12", "o17", "o22", "o25", "o28", "o30"
)

# assign colours to variables
colour.osiq = as.vector(rep("", ncol(df.osiq)))

for(i in 1:ncol(df.osiq)) {
  switch(substr(names(df.osiq)[i], 1, 1),
         "o" = {colour.osiq[i] = colourCode.osiq[1]},
         "s" = {colour.osiq[i] = colourCode.osiq[2]})
}
```

```{r plsc_sam_osiq}
plsc.sam.osiq = tepPLS(
  DATA1 = df.osiq,
  DATA2 = df.sam,
  center1 = TRUE, scale1 = FALSE,
  center2 = TRUE, scale2 = FALSE,
  make_design_nominal = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
plsc.sam.osiq$TExPosition.Data$t
```

<!-- ======================================================================= -->

# Supplemental analyses

## Factor structure of the SAM in two independent samples

### 2013 dataset

```{r setup_sam_og}
# load SAM data from independent sample
df.og = read_excel("../../../../damp/SAMOriginalData_FinalSampleNof598.xls")

# select SAM items
df.og.sam = df.og[ , which(colnames(df.og) == 'E1') : which(colnames(df.og) == 'Sp20')]

# IDs and n for subtables
domainNames = wrapr::qc(Episodic, Future, Semantic, Spatial) 
domainNamesVar = wrapr::qc(Episodic, Future, Semantic, sPatial) 
posUp = sapply(strsplit(domainNamesVar, ''), function(a) which(a %in% LETTERS)[1])
domainKeys = substr(domainNamesVar, posUp, posUp)
domainNs = c(42, 16, 24, 20)
domains = data.frame(domainNames, domainNamesVar, domainKeys, domainNs, colourCode.sam,
                     row.names = domainNames,
                     stringsAsFactors = FALSE)

# rename variables
colnames(df.og.sam) = paste0(rep(domains$domainKeys, domains$domainNs),
                             unlist(sapply(domains$domainNs,function(x){1:x})))

# keep only items from final published questionnaire
keepSEpisodic = c(9, 10, 13, 21, 28, 35, 37, 39)
keepSFuture = c(1, 3, 5, 7, 2, 15)
keepSSemantic = c(4, 5, 11, 12, 13, 22)
keepSSpatial = c(7, 3, 5, 1, 12, 17)
keepMini = c(keepSEpisodic, 
             42 + keepSFuture, 
             42 + 16 + keepSSemantic, 
             42 + 16 + 24 + keepSSpatial)
df.og.sam = df.og.sam[ , keepMini]

# rename variables
names(df.og.sam) = c(
  paste0("E", seq(1, 8)),
  paste0("F", seq(1, 6)),
  paste0("M", seq(1, 6)),
  paste0("S", seq(1, 6))
)
```

```{r pca_sam_og}
pca.sam.og = epPCA(
  df.og.sam,
  scale = FALSE,
  center = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
pca.sam.og$ExPosition.Data$t
```

### 2019 dataset

```{r demographics_et}
# load SAM data from independent sample
df.et = read.csv("../../../baytourET/data/baytourET_samscreen/2020-06-23-baytourET-samscreen-clean.csv")

# number of participants
nrow(df.et)

# age 
mean(df.et$age)
sd(df.et$age)
range(df.et$age)

# education
mean(df.et$edu)
sd(df.et$edu)
range(df.et$edu)

# gender
df.et %>% 
  group_by(gender) %>% 
  summarise(n  = n())
```

```{r setup_sam_et}
# select SAM items
df.sam.et = df.et %>% 
  select(starts_with("sam"), -c(sam_epi, sam_sem, sam_spa, sam_fut))

# rename variables
names(df.sam.et) = c(
  paste0("E", seq(1, 8)),
  paste0("F", seq(1, 6)),
  paste0("M", seq(1, 6)),
  paste0("S", seq(1, 6))
)
```

```{r pca_sam_et}
pca.sam.et = epPCA(
  df.sam.et,
  scale = FALSE,
  center = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
pca.sam.et$ExPosition.Data$t
```

## Relationship betwen SAM and OSIQ

### Replication in an independent sample (2019 dataset)

#### Univariate analysis (linear regression)

```{r setup_sam_osiq_et}
# select SAM and OSIQ items
df.sam.osiq.et = df.et %>%
  select(starts_with("sam"),
         osiq_s_1, osiq_s_9, osiq_s_13, osiq_s_14, osiq_s_18, osiq_s_27r, osiq_s_29,
         osiq_o_4, osiq_o_12, osiq_o_17, osiq_o_22, osiq_o_25, osiq_o_28, osiq_o_30)

## convert to numeric
df.sam.osiq.et %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

# calculate OSIQ domain scores based on mini OSIQ items

## object
df.sam.osiq.et$osiq_mini_o =
  df.sam.osiq.et %>%
  select(starts_with("osiq_o")) %>%
  rowMeans()

## spatial
df.sam.osiq.et$osiq_mini_s =
  df.sam.osiq.et %>%
  select(starts_with("osiq_s")) %>%
  rowMeans()

# calculate SAM domain scores based on raw item scores

## episodic
df.sam.osiq.et$sam_epi_raw = df.sam.osiq.et %>%
  select(starts_with("sam_event")) %>%
  rowMeans()

## spatial
df.sam.osiq.et$sam_spa_raw = df.sam.osiq.et %>%
  select(starts_with("sam_spatial")) %>%
  rowMeans()

## semantic
df.sam.osiq.et$sam_sem_raw = df.sam.osiq.et %>%
  select(starts_with("sam_semantic")) %>%
  rowMeans()

## future
df.sam.osiq.et$sam_fut_raw = df.sam.osiq.et %>%
  select(starts_with("sam_future")) %>%
  rowMeans()
```

```{r regression_sam_osiq.et}
# SAM-episodic
model_samE_osiq.et = lm(sam_epi_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq.et)
model_samE_osiq.et %>% summary()
model_samE_osiq.et %>% modelEffectSizes()

# SAM-spatial
model_samS_osiq.et = lm(sam_spa_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq.et)
model_samS_osiq.et %>% summary()
model_samS_osiq.et %>% modelEffectSizes()

# SAM-future
model_samF_osiq.et = lm(sam_fut_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq.et)
model_samF_osiq.et %>% summary()
model_samF_osiq.et %>% modelEffectSizes()

# SAM-semantic
model_samM_osiq.et = lm(sam_sem_raw ~ osiq_mini_o + osiq_mini_s, data = df.sam.osiq.et)
model_samM_osiq.et %>% summary()
model_samM_osiq.et %>% modelEffectSizes()
```

#### Multivariate analysis (partial least squares correlation)

```{r setup_osiq_et}
# select mini OSIQ items
df.osiq.et = df.et %>%
  select(osiq_s_1, osiq_s_9, osiq_s_13, osiq_s_14, osiq_s_18, osiq_s_27r, osiq_s_29,
         osiq_o_4, osiq_o_12, osiq_o_17, osiq_o_22, osiq_o_25, osiq_o_28, osiq_o_30)

## convert to numeric
df.osiq.et %<>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

# rename variables
names(df.osiq.et) = c(
  "s1", "s9", "s13", "s14", "s18", "s27", "s29",
  "o4", "o12", "o17", "o22", "o25", "o28", "o30"
)
```

```{r plsc_sam_osiq_et}
plsc.sam.osiq.et = tepPLS(
  DATA1 = df.osiq.et,
  DATA2 = df.sam.et,
  center1 = TRUE, scale1 = FALSE,
  center2 = TRUE, scale2 = FALSE,
  make_design_nominal = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
plsc.sam.osiq.et$TExPosition.Data$t
```

## Gender differences

```{r pca_sam_female}
pca.sam.female = epPCA(
  df.sam.female,
  scale = FALSE,
  center = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
pca.sam.female$ExPosition.Data$t
```

```{r pca_sam_male}
pca.sam.male = epPCA(
  df.sam.male,
  scale = FALSE,
  center = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
pca.sam.male$ExPosition.Data$t
```

```{r regression_sam_osiq_gender}
# set up data
df.sam.osiq.gender = cbind(df$gender, df.sam.osiq)
names(df.sam.osiq.gender)[1] = "gender"
df.sam.osiq.gender %<>% filter(gender != "Prefer not to answer")

# SAM-episodic
model_samE_osiq_gender = 
  lm(sam_epi_raw ~ gender * (osiq_mini_o + osiq_mini_s), data = df.sam.osiq.gender)
summary(model_samE_osiq_gender)$adj.r.squared
model_samE_osiq_gender %>% car::Anova(type = 3)
model_samE_osiq_gender %>% modelEffectSizes()

# SAM-spatial
model_samS_osiq_gender = 
  lm(sam_spa_raw ~ gender * (osiq_mini_o + osiq_mini_s), data = df.sam.osiq.gender)
summary(model_samS_osiq_gender)$adj.r.squared
model_samS_osiq_gender %>% car::Anova(type = 3)
model_samS_osiq_gender %>% modelEffectSizes()

# SAM-future
model_samF_osiq_gender = 
  lm(sam_fut_raw ~ gender * (osiq_mini_o + osiq_mini_s), data = df.sam.osiq.gender)
summary(model_samF_osiq_gender)$adj.r.squared
model_samF_osiq_gender %>% car::Anova(type = 3)
model_samF_osiq_gender %>% modelEffectSizes()

# SAM-semantic
model_samM_osiq_gender = 
  lm(sam_sem_raw ~ gender * (osiq_mini_o + osiq_mini_s), data = df.sam.osiq.gender)
summary(model_samM_osiq_gender)$adj.r.squared
model_samM_osiq_gender %>% car::Anova(type = 3)
model_samM_osiq_gender %>% modelEffectSizes()
```

## Personality effects

### Regressions of OSIQ on SAM

```{r calculate_halo}
# select items to analyze with halo
df.halo = df %>% 
  select(starts_with("phq"),
         starts_with("bfi"))

# change items from factor to numeric
df.halo %<>%
  mutate_all(as.character) %>% 
  mutate_all(as.numeric)

# add OSIQ and SAM scores to halo dataframe
df.halo = cbind(df.halo, df.sam.osiq)

# standardize BFI domain scores
df.halo %<>% 
  mutate_at(.vars = c("bfi_e", "bfi_a", "bfi_c", "bfi_n", "bfi_o"),
            scale)

# calculate halo scores
df.halo$halo = 
  ((df.halo$bfi_a + 
    df.halo$bfi_c + 
    df.halo$bfi_o + 
    df.halo$bfi_e) / 4) -
  df.halo$bfi_n
```

```{r halo_osiq_sam}
# episodic
model_samE_osiq_halo =
  lm(sam_epi_raw ~ halo + osiq_mini_o + osiq_mini_s, data = df.halo)
summary(model_samE_osiq_halo)
modelEffectSizes(model_samE_osiq_halo)

# spatial
model_samS_osiq_halo =
  lm(sam_spa_raw ~ halo + osiq_mini_o + osiq_mini_s, data = df.halo)
summary(model_samS_osiq_halo)
modelEffectSizes(model_samS_osiq_halo)

# semantic
model_samM_osiq_halo =
  lm(sam_sem_raw ~ halo + osiq_mini_o + osiq_mini_s, data = df.halo)
summary(model_samM_osiq_halo)
modelEffectSizes(model_samM_osiq_halo)

# future
model_samF_osiq_halo =
  lm(sam_fut_raw ~ halo + osiq_mini_o + osiq_mini_s, data = df.halo)
summary(model_samF_osiq_halo)
modelEffectSizes(model_samF_osiq_halo)
```

### BFI and SAM

```{r setup_bfi}
# select BFI items
df.bfi = df %>% 
  select(starts_with("bfi")) %>% 
  select(-c("bfi_e", "bfi_a", "bfi_c", "bfi_n", "bfi_o"))

# convert to numeric
df.bfi %<>% 
  mutate_all(as.character) %>% 
  mutate_all(as.numeric)

# rename variables
for (i in 1:length(names(df.bfi))) {
  
  # remove "bfi_" from beginning of column names
  names(df.bfi)[i] %<>% substr(5, nchar(names(df.bfi)[i]))
  
  # remove "r" from end of reverse coded items 
  if (substr(names(df.bfi)[i], 
             nchar(names(df.bfi)[i]), 
             nchar(names(df.bfi)[i])) == "r") {
    names(df.bfi)[i] %<>% substr(1, nchar(names(df.bfi)[i])-1)
  }
  
  # remove underscore
  names(df.bfi)[i] = gsub(pattern = "_", replacement = "", x = names(df.bfi)[i])
  
}

# assign colours to variables
colour.bfi = as.vector(rep("", ncol(df.bfi)))

for (i in 1:ncol(df.bfi)) {
  switch(substr(names(df.bfi)[i], 1, 1),
         a = {colour.bfi[i] = colourCode.bfi[1]},
         c = {colour.bfi[i] = colourCode.bfi[2]},
         e = {colour.bfi[i] = colourCode.bfi[3]},
         n = {colour.bfi[i] = colourCode.bfi[4]},
         o = {colour.bfi[i] = colourCode.bfi[5]})
}
```

```{r plsc_sam_bfi}
plsc.sam.bfi = tepPLS(
  DATA1 = df.sam, 
  DATA2 = df.bfi,
  center1 = TRUE, scale1 = FALSE,
  center2 = TRUE, scale2 = FALSE,
  make_design_nominal = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
plsc.sam.bfi$TExPosition.Data$t
```

## Depression effects

```{r setup_phq}
# select PHQ items
df.phq = df %>% 
  select(starts_with("phq")) %>% 
  select(-c("phq_function", "phq_total"))

# convert to numeric
df.phq %<>% 
  mutate_all(as.character) %>% 
  mutate_all(as.numeric)

# rename variables
for (i in 1:length(names(df.phq))) {
  
  # remove underscore
  names(df.phq)[i] = gsub(pattern = "_", replacement = "", x = names(df.phq)[i])
  
}
```

```{r plsc_sam_phq}
plsc.sam.phq = tepPLS(
  DATA1 = df.sam, 
  DATA2 = df.phq,
  center1 = TRUE, scale1 = FALSE,
  center2 = TRUE, scale2 = FALSE,
  make_design_nominal = TRUE,
  graphs = FALSE)

# percent variance explained by each dimension
plsc.sam.phq$TExPosition.Data$t
```


<!-- ======================================================================= -->


# Export workspace

```{r export_workspace}
save.image("analysis_workspace.RData")
```

<!-- ======================================================================= -->


# Session info

```{r session_info}
sessionInfo()
```
